\documentclass[12pt]{article}
\usepackage[margin=0.5in]{geometry}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{silence}
\usepackage{mathrsfs}
\WarningsOff[latex]
\usepackage{float}
\graphicspath{ {./pics/} }
\linespread{1.2}
\newcommand{\code}{\texttt}

\begin{document}
\subsection*{Exercise 1.}
\paragraph*{(a)(i)} We prove that \(\delta\) is well defined by a contrapositive proof i.e. if \(\exists a \in \Sigma \ [xa] \neq [ya]\), then \([x] \neq [y]\).
Suppose \([xa] \neq [ya]\) for \(a \in \Sigma\), then \((xa, ya) \notin R_{L}\). This means that \(\exists w \in \Sigma^{*}\) such that \(xaw \in L\) but \(yaw \notin L\) or vice versa. Now let \(z = aw\), we have \(xz \in L \land yz \notin L\) or vice versa. Hence, \((x, y) \notin R_{L}\) and \([x] \neq [y]\). We conclude by contrapositive that if \([x] = [y]\) then \([xa] = [ya]\) for all \(a \in \Sigma\).
\paragraph*{(ii)} We prove a more general statement i.e. \(\forall x \in \Sigma^{*}\ \hat{\delta}([\epsilon], x) = [x]\).\\
This implies \(\hat{\delta}([\epsilon], x) \in F\) iff \(x \in L\). Since if \(x \in L\), then by the definition of \(F\), \([x] = \hat{\delta}([\epsilon], x) \in F\). Conversely, if \(\hat{\delta}([\epsilon], x) \in F \Leftrightarrow [x] \in F\), then by the definition of \(F\), \(x \in L\). \\
Now we prove the lemma by an induction on \(|x|\).
\begin{itemize}
  \item \textbf{Basis case} \(x = \epsilon\). \(\delta([\epsilon], \epsilon) = [\epsilon]\) by the definition of the transition function.
  \item \textbf{Step case} Assume that \(\hat{\delta}([\epsilon], x) = [x]\) for \(|x| < k\). We prove for \(w = xa\) where \(x \in \Sigma^{k-1}, a \in \Sigma\).
  \begin{align*}
    \hat{\delta}([\epsilon], xa) &= \delta(\hat{\delta}([\epsilon], x), a) \tag*{(definition of \(\hat{\delta}\))} \\
    &= \delta([x], a) \tag*{(IH)} \\
    &= [xa] \tag*{(definition of \(\delta\))}
  \end{align*}
\end{itemize}
Hence, we have proven the lemma.

\noindent (b). Denote the DFA in part (a) as \(D = (Q, \Sigma, \delta, q_{0}, F)\). Let DFA \(A = (Q_{A}, \Sigma, \delta_{A}, q_{0_{A}}, F_{A})\) with \(L(A) = L\) and no unreachable state. Construct \(f : Q_A \to Q\)
\begin{equation}
  f(\hat{\delta}_A(q_{0_{A}}, w)) = \hat{\delta}(q_0, w) \text{ for } \forall w \in \Sigma^*
\end{equation}
\(f\) is well defined since
\begin{itemize}
  \item for each \(q \in Q_{A}\), there exists a \(f(q) \in Q\). Let \(q \in Q_A\), then \(\exists w \in \Sigma^*\) such that \(q = \hat{\delta}_A(q_{0_{A}}, w)\) since \(q\) is a reachable state by construction. By def of the constructed \(f\), we have \(f(q) = \hat{\delta}(q_0, w) = \hat{\delta}([\epsilon], w) \overset{part (b)}{=} [w] \in Q\)
  \item and each \(q \in Q_{A}\) has a unique mapping \(f(q) \in Q\). Suppose \(\hat{\delta}_A(q_{0_{A}}, x) = \hat{\delta}_A(q_{0_{A}}, y)\) for \(x \neq y\), we show \(\hat{\delta}(q_0, x) = \hat{\delta}(q_0, y)\). Let \(w \in \Sigma^*\). \(xw \in L \Leftrightarrow \hat{\delta}_A(q_{0_{A}}, xw) \in F_A \Leftrightarrow \hat{\delta}_A(\hat{\delta}_A(q_{0_{A}}, x), w) \in F_A \Leftrightarrow \hat{\delta}_A(\hat{\delta}_A(q_{0_{A}}, y), w) \in F_A \Leftrightarrow \hat{\delta}_A(q_{0_{A}}, yw) \in F_A \Leftrightarrow yw \in L\). Hence, \(\forall w \in \Sigma^*, xw \in L \Leftrightarrow yw \in L\), i.e. \((x,y) \in R_L \Leftrightarrow [x] = [y] \Leftrightarrow \hat{\delta}(q_0, x) = \hat{\delta}(q_0, y)\)
\end{itemize}
If \(A\) has fewer states than \(D\), then \(f\) means there exists at least one state \([x] \in D\) such that for \(\forall w \in \Sigma^*\), \(f(\hat{\delta}_A(q_{0_{A}}, w)) \neq [x]\). However, \(f(\hat{\delta}_A(q_{0_{A}}, x)) = \hat{\delta}(q_0, x) = [x]\), this is a contradiction. Therefore, \(A\) must have at least as many states as \(D\). For \(A\) with unreachable states, it must be more. Hence \(D\) is the minimal DFA for \(L\). 

We prove that any DFA \(A = (Q_{A}, \Sigma, \delta_{A}, q_{0_{A}}, F_{A})\) with \(L(A) = L\) has at least as many states as \(D\), by constructing a surjective function \(f\) from \(Q_{A}\) to \(Q\). We adapted the proof from solutions of Tutorial 2. Define \(S(q) = \{w \in \Sigma^{*} \mid \hat{\delta}(q_{0_{A}}, w) = q\}\), and \(f\) 
\begin{align*}
  &f : \{q \in Q_{A} \mid S(q) \neq \emptyset\} \to \{[x] \mid x \in \Sigma^{*}\} \\
  &f(q) = [x] \text{ with } S(q) \subseteq [x]
\end{align*}
We show that \(f\) is well defined, i.e.
\begin{itemize}
  \item \(\forall q \in Q_{A}\) such that \(S(q) \neq \emptyset\), \(\exists [x] \in Q\) with \(S(q) \subseteq [x]\)\\
  Proof. For \(q \in Q_{A}\), pick \([x] \in Q\) such that \(x \in S(q)\). Now for \(\forall u, v \in S(q)\), we proved in the tutorial that \((u, v) \in R_{L}\). Hence for \(\forall u \in S(q)\), \((x, u) \in R_{L}\). Hence, \(u \in [x]\), and \(S(q) \subseteq [x]\)
  \item if \(f(q) = [x]\) and \(f(q) = [y]\) for \(x \neq y\), then \([x] = [y]\)\\
  Proof. Suppose \(f(q) = [x]\) and \(f(q) = [y]\), then \(S(q) \subseteq [x]\) and \(S(q) \subseteq [y]\). Then \(u \in S(q) \Rightarrow u \in [x] \Leftrightarrow (x, u) \in R_{L}\); and \(u \in S(q) \Rightarrow u \in [y] \Leftrightarrow (y, u) \in R_{L} \Leftrightarrow (u, y) \in R_{L}\). By transitivity, \((x, u) \in R_{L} \land (u, y) \in R_{L} \Rightarrow (x, y) \in R_{L} \Leftrightarrow [x] = [y]\)
\end{itemize}
Now, we show that \(f\) is surjective. For \([x] \in Q\), we pick \(q = \hat{\delta}(q_{0_{A}}, x) \in Q_{A}\). \(S(q) \neq \emptyset\) since \(x \in S(q)\) by definition. We only need to show \(S(q) \subseteq [x]\). By the same argument that \(\forall u, v \in S(q), (u, v) \in R_{L}\). \(u \in S(q) \Rightarrow (x, u) \in R_{L} \Leftrightarrow u \in [x]\).\\ 
Since \(f\) is surjective, then for every state \([x]\) in DFA \(D\), there are corresponding reachable states in DFA \(A\) that are mapped to \([x]\), and the cardinality of \(Q_{A}\) exclusive of unreachable states is at least the cardinality of \(Q\). Hence, any DFA that accepts \(L\) would have at least as many states as DFA \(D\). Then \(D\) is the minimal DFA.

\subsection*{Exercise 2.} 
\paragraph*{(a)} Pick n=7. We consider \(w=a^jc^ka^lb^m \in L\) with \(|w| \geq 7\) in two separate cases.
\begin{itemize}
  \item \(0 \leq j \leq 5\), \(k \geq 2\) and \(l \neq m\)\\
  Divide \(w = xyz\) as \(x = a^jc\), \(y=c\), and \(z=c^{k-2}a^lb^m\). \(|xy| = j+2 \leq 7\) since \(0 \leq j \leq 5\), \(|y| = 1 > 0\), and by pumping \(y\), we get
  \begin{itemize}
    \item \(xy^0z = a^jcc^{k-2}a^lb^m = a^jc^{k-1}a^lb^m\)\\
    if \(k = 2\), then \(k-1 = 1\), \(l \neq m\) still means \(xy^0z \in L\)\\
    if \(k > 2\), then \(k-1 \geq 2\), \(l \neq m\) still holds, and \(xy^0z \in L\)
    \item \(xy^iz = a^jcc^ic^{k-2}a^lb^m = a^jc^{i+k-1}a^lb^m\) for \(i > 1\)\\
    \(i+k-1 > 2\) for \(k \geq 2\), \(l \neq m\) still holds, so \(xy^iz \in L\) for \(i > 1\)
  \end{itemize}
  Hence, all \(w\) when \(k \geq 2\) satisfies the pumping lemma.
  \item \(0 \leq j \leq 5\), \(k < 2\) and \(k, l, m \in \mathbb{N}\)\\
  Note that for \(w \in L\) with \(|w| \geq 7\), \(w\) should have at least one of \(l\) and \(m\) not being zero since \(j+k < 7\)
  \begin{itemize}
    \item if \(l \neq 0 \Leftrightarrow l \geq 1\), then we divide \(w = xyz\) such that \(x=a^jc^k\), \(y=a\), \(z=a^{l-1}b^m\). \(|xy|= j+k+1 \leq 7\), \(|y| = 1>0\), and \(xy^iz = a^jc^ka^ia^{l-1}b^m = a^jc^ka^{i+l-1}b^m\) and \(i+l-1 \geq 0\) for \(i \in \mathbb{N}\), hence \(xy^iz \in L\) for \(\forall i \in \mathbb{N}\)
    \item if \(l = 0\), then \(m \neq 0 \Leftrightarrow m \geq 1\). We divide \(w = xyz\) such that \(x = a^jc^ka^0\), \(y = b\), \(z = b^{m-1}\). \(|xy|= j+k+1 \leq 7\), \(|y| = 1>0\), and \(xy^iz = a^jc^ka^0b^ib^{m-1}= a^jc^ka^0b^{i+m-1}\) and \(i+m-1 \geq 0\) for \(i \in \mathbb{N}\), hence \(xy^iz \in L\) for \(\forall i \in \mathbb{N}\)
  \end{itemize}
\end{itemize}
Hence, we have proven \(L\) satisfies the pumping lemma.

\paragraph*{(b)} We show that \(R_L\) has an infinite number of equivalence classes. \\
Consider \(u = a^jc^ka^lb^m\) and \(v = a^jc^ka^{l'}b^{m'}\) where \(0 \leq j \leq 5\), \(k \geq 2\), \(l > m\) and \(l' > m'\), \(l - m \neq l' - m'\). Since \(k \geq 2 \land l \neq m \land l' \neq m'\), by definition \(u, v \in L\). Now pick \(w = b^{l-m}\), \(uw = a^jc^ka^lb^mb^{l-m} = a^jc^ka^lb^l \notin L\), but \(vw = a^jc^ka^{l'}b^{m'}b^{l-m} = a^jc^ka^{l'}b^{m'+l-m} \in L\) since \(m'+l-m \neq l'\). Hence, \((u, v) \notin R_L \Leftrightarrow [u] \neq [v]\). In other words, for any pair of \((u, v) \in L\) with \(k \geq 2\) and distinct positive values of \(l-m\), we have \([u] \neq [v]\). Since there is an infinite number of distinct positive values of \(l-m\) for \(l, m \in \mathbb{N}\), there is an infinite number of distinct equivalence classes in \(R_L\). By the Myhill-Nerode Theorem, \(L\) is not regular.

\subsection*{Exercise 3.}
\paragraph*{(a)} We prove a lemma i.e. \(f(\hat{\delta}(q, w)) = \hat{\delta}'(f(q), w)\) for \(\forall w \in \Sigma^*\), \(\forall q \in Q\).\\
Let \(q \in Q\), by an induction on \(|w|\),
\begin{itemize}
  \item \textbf{Base case} \(w = \epsilon\)\\
  \(f(\hat{\delta}(q, \epsilon))\) = \(f(q)\) by the def of \(\hat{\delta}\) and \(\hat{\delta}'(f(q), \epsilon) = f(q)\) by def of \(\hat{\delta}'\), hence \(f(\hat{\delta}(q, \epsilon)) = \hat{\delta}'(f(q), \epsilon)\)
  \item \textbf{Step case} Assume the claim for \(\forall w\) with \(|w| < n\). Now we prove for \(w = xa\) where \(x \in \Sigma^{n-1}\), \(a \in \Sigma\)
  \begin{align*}
    f(\hat{\delta}(q, xa)) &= f(\delta(\hat{\delta}(q, x), a)) \tag*{(def of \(\hat{\delta})\)}\\
    &= \delta'(f(\hat{\delta}(q, x)), a) \tag*{(def (3) of \(f\))}\\
    &= \delta'(\hat{\delta}'(f(q), x), a) \tag*{(IH since \(|x| < n\))}\\
    &= \hat{\delta}'(f(q), xa) \tag*{(def of \(\hat{\delta}'\))}
  \end{align*}
\end{itemize}
Hence, we have proven the lemma.\\
Now we prove \(\mathscr{L}(P, q) = \mathscr{L}(P', f(q))\) for \(\forall q \in Q\). Let \(q \in Q\), \(w \in \Sigma^*\)
\begin{align*}
  w \in \mathscr{L}(P, q) &\Leftrightarrow \hat{\delta}(q, w) \in F \\ 
  &\Leftrightarrow f(\hat{\delta}(q, w)) \in F' \tag*{(def (2) of \(f\))}\\
  &\Leftrightarrow \hat{\delta}'(f(q), w) \in F' \tag*{(lemma)} \\
  &\Leftrightarrow w \in \mathscr{L}(P', f(q))
\end{align*}
Hence the claim.

\paragraph*{(b)} 
\begin{align*}
  L(P) &= \{w \in \Sigma^* \mid \hat{\delta}(q_0, w) \in F\}\\
  &= \mathscr{L}(P, q_0) \tag*{(def of \(\mathscr{L}\))}\\
  &= \mathscr{L}(P', f(q_0) \tag*{(part (a))}\\
  &= \mathscr{L}(P', q'_0) \tag*{(def (1) of \(f\))}\\
  &= \{w \in \Sigma^* \mid \hat{\delta}'(q'_0, w) \in F'\} \tag*{(def of \(\mathscr{L}\))}\\
  &= L(P')
\end{align*}

\subsection*{Exercise 4.}
\paragraph*{1.} We show validity of \((r^*)^* = r^*\) by proving \(L((r^*)^*) = L(r^*)\).\\
\(L((r^*)^*) = L(r^*)^* = (L(r)^*)^* = L(r)^* = L(r^*)\) where the third equaltiy uses the algebraic law of Kleene-*. Now we prove for \(\forall L \subseteq \Sigma^*, (L^*)^* = L^*\). Let \(L \subseteq \Sigma^*\),
\begin{itemize}
  \item \(L^* \subseteq (L^*)^*\) since \(L^* = (L^*)^1 \subseteq (L^*)^*\)
  \item \((L^*)^* \subseteq L^*\)\\
  Let \(w \in (L^*)^*\), we can write \(w = w_1w_2...w_n\) for \(n \geq 0\) where each \(w_i \in L^*\). We can also write each \(w_i = x_{i1}x_{i2}...x_{il_i}\) for \(l_i \geq 0\) where each \(x_{il_i} \in L\). Then \(w = x_{11}x_{12}...x_{1l_1}..x_{n1}x_{n2}...x_{nl_n} \in L^{\sum_1^n{l_n}} \subseteq L^*\). Hence, \(w \in L^*\). 
\end{itemize}

\paragraph*{2.} We prove \(L((r+s)^*) = L((r^*s)^*r^*)\). Denote \(R = L(r)\) and \(S = L(s)\) for clarity. By def of reg lang,\\
\(L((r+s)^*) = L(r+s)^* = (L(r) \cup L(s))^* = (R \cup S)^*\)\\
\(L((r^*s)^*r^*) =  L((r^*s)^*)L(r^*) = L(r^*s)^*L(r)^* = (L(r^*)L(s))^*R^* = (L(r)^*S)^*R^* = (R^*S)^*R^*\)\\
In other words, we prove \((R \cup S)^* = (R^*S)^*R^*\).
\begin{itemize}
  \item \((R \cup S)^* \subseteq (R^*S)^*R^*\)\\
  We prove the lemma that for \(\forall n \geq 0\), \((R \cup S)^n \subseteq (R^*S)^*R^*\), then by def of set union, \((R \cup S)^* = \cup_{n \geq 0}(R \cup S)^n \subseteq (R^*S)^*R^*\). By an induction on n,
  \begin{itemize}
    \item n = 0 \((R \cup S)^0 = \{\epsilon\} = \{\epsilon\}\{\epsilon\} = (R^*S)^0R^0 \subseteq (R^*S)^*R^*\)
    \item prove \((R \cup S)^{n+1} \subseteq (R^*S)^*R^*\)
    \begin{align*}
      (R \cup S)^{n+1} &= (R \cup S)^{n} (R \cup S) \tag*{(concat of lang)}\\
      &= \{xa \in \Sigma^* \mid x \in (R \cup S)^{n}, a \in (R \cup S)\} \tag*{(set notation of concat)}\\
      &\subseteq \{xa \in \Sigma^* \mid x \in (R^*S)^*R^*, a \in (R \cup S)\} \tag*{(IH)}\\
      &= \{xa \in \Sigma^* \mid x \in (R^*S)^*R^*, a \in R\} \cup \{xa \in \Sigma^* \mid x \in (R^*S)^*R^*, a \in S\} \\
      &= ((R^*S)^*R^*)R \cup ((R^*S)^*R^*)S \tag*{(set def of concat)}
    \end{align*}
    We show both subsets are in \((R^*S)^*R^*\). But first we prove some laws.\\
    \textbf{L1.} \(\forall L \subseteq \Sigma^*\), \(L^*L \subseteq L^*\)\\
    Proof. \(L^*L = (\cup_{n \geq 0}L^n)L = \{xa \in \Sigma^* \mid x \in \cup_{n \geq 0}L^n, a \in L\} = \cup_{n \geq 0}\{xa \in \Sigma^* \mid x \in L^n, a \in L\} = \cup_{n \geq 0}(L^nL) = \cup_{n \geq 0}L^{n+1} = \cup_{n \geq 1}L^{n} \subseteq L^*\)\\
    \textbf{L2.} \(\forall L, M, N \subseteq \Sigma^*\), \(M \subseteq N \Rightarrow LM \subseteq LN\)\\
    Proof. Suppose \(M, N \subseteq \Sigma^*\) with \(M \subseteq N\). Let \(w \in LM\). Write \(w = xa\), where \(x \in L, a \in M\). Since \(M \subseteq N\), \(a \in N\). Then \(w = xa \in LN\). In short, \(LM = \{xa \in \Sigma^* \mid x \in L, a \in M\} \subseteq \{xa \in \Sigma^* \mid x \in L, a \in N\} = LN\)\\
    Back to the proof
    \begin{itemize}
      \item \(((R^*S)^*R^*)R \subseteq (R^*S)^*R^*\)
      \begin{align*}
        &R^*R \subseteq R^* \tag*{(L1)}\\
        &\Rightarrow (R^*S)^*(R^*R) \subseteq (R^*S)^*R^* \tag*{(L2)}\\
        &\Leftrightarrow ((R^*S)^*R^*)R \subseteq (R^*S)^*R^* \tag*{(associativity)}
      \end{align*}
      \item \(((R^*S)^*R^*)S \subseteq (R^*S)^*R^*\)
      \begin{align*}
        & (R^*S)^*(R^*S) \subseteq (R^*S)^* \tag*{(L1)}\\
        &\Leftrightarrow ((R^*S)^*R^*)S \subseteq (R^*S)^* = (R^*S)^*\{\epsilon\}
        = (R^*S)^*R^0 \subseteq (R^*S)^*R^*
        \tag*{(assoc \& concat with \(\{\epsilon\}\))}
      \end{align*}
    \end{itemize}
    Hence the step case. 
  \end{itemize}
  Hence the lemma.
  \item \((R^*S)^*R^* \subseteq (R \cup S)^*\)\\
  \textbf{L3.} for \(\forall M,N \in \Sigma^*\), \(M \subseteq N \Rightarrow M^* \subseteq N^*\)\\
  Proof. Suppose \(M \subseteq N\). Let \(w \in M^*\), then we can write \(w = w_1w_2...w_n\) where each \(w_i 
  \in M\). Since \(M \subseteq N\), hence each \(w_i \in N\). Then \(w = w_1w_2...w_n \in N^*\). Hence, \(M^* \subseteq N^*\).\\
  \textbf{L4.} for \(\forall M,N,X,Y \in \Sigma^*\), \(M \subseteq N \land X \subseteq Y \Rightarrow MX \subseteq NY\)\\
  Proof. Suppose \(M \subseteq N \land X \subseteq Y\). Let \(w \in MX\), then \(w = ax\) where \(a \in M\), \(x \in X\). Since \(M \subseteq N\) and \(X \subseteq Y\), then \(a \in N\) and \(x \in Y\). Hence, \(w \in NY\). \(MX \subseteq NY\).\\
  Now we can prove \((R^*S)^*R^* \subseteq (R \cup S)^*\).\\
  Since \(R \subseteq (R \cup S)\), by L3 we have \(R^* \subseteq (R \cup S)^*\). Also, \(S \subseteq (R \cup S)\), by L4 \(R^*S \subseteq (R \cup S)^*(R \cup S) \overset{L1}{\subseteq} (R \cup S)^*\). By L3 again, \((R^*S)^* \subseteq ((R \cup S)^*)^* \overset{part 1}{=} (R \cup S)^*\). By L4, \((R^*S)^*R^* \subseteq (R \cup S)^*(R \cup S)^* = ((R \cup S)^*)^2 \subseteq ((R \cup S)^*)^* \overset{part 1}{=} (R \cup S)^*\).
\end{itemize}
Hence, we have formally proven \(L((r+s)^*) = L((r^*s)^*r^*)\), thus the validity of \((r+s)^* = (r^*s)^*r^*\).

\paragraph*{3.} Let \(R = L(r)\), \(S = L(s)\).\\
\(L((rs)^*) = L(rs)^* = (L(r)L(s))^* = (RS)^*\)\\
\(L(\epsilon + r(sr)^*s) = L(\epsilon) \cup L(r(sr)^*s) = \{\epsilon\} \cup L(r(sr)^*)L(s) = \{\epsilon\} \cup L(r)L((sr)^*)S = \{\epsilon\} \cup RL(sr)^*S = \{\epsilon\} \cup R(L(s)L(r))^*S = \{\epsilon\} \cup R(SR)^*S\)\\
We prove \((RS)^* = \{\epsilon\} \cup R(SR)^*S\) by equaltiy of the subsets, i.e, \((RS)^0 = \{\epsilon\}\) and \((RS)^n = R(SR)^{n-1}S\) for \(\forall n \geq 1\).
\begin{itemize}
  \item \((RS)^0 = \{\epsilon\}\) by definition
  \item we prove by induction on \(n\) that \((RS)^n = R(SR)^{n-1}S\) for \(\forall n \geq 1\).
  \begin{itemize}
    \item \textbf{Base case} n = 1. \((RS)^1 = RS = (R\{\epsilon\})S = (R(SR)^0)S = R(SR)^0S \)
    \item \textbf{Step case} \((RS)^{n+1} = (RS)^n(RS) \overset{IH}{=} (R(SR)^{n-1}S)(RS) \overset{assoc}{=} ((R(SR)^{n-1}S)R)S\\ \overset{assoc}{=} (R(SR)^{n-1}(SR))S = (R(SR)^{n})S = R(SR)^{n}S\)
  \end{itemize}
  Hence \(\cup_{n \geq 1}(RS)^n = \cup_{n \geq 1} R(SR)^{n-1}S = \cup_{n \geq 0} R(SR)^{n}S\)
\end{itemize}
From above, \((RS)^* = \cup_{n \geq 0}(RS)^n = (RS)^0 \cup (\cup_{n \geq 1}(RS)^n)\) = \(\{\epsilon\} \cup (\cup_{n \geq 0} R(SR)^{n}S) = \{\epsilon\} \cup R(SR)^*S\). Hence, \((rs)^* = \epsilon + r(sr)^*s\) is valid.
\end{document}